Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-10-17 paper/sections/proof-reconstruction.tex
[ paper ] minor fixies.

In the rest of this section, we provide a formal description using
type theory to build definitions and theorems for functions
necessary to emulate \Metis' inference rules. For this description,
we will write the type judgments with ``$a : A$'' to denote that $a$
has $\abbre{A}$ type. Types are written using small caps.
We use \Prop type for propositions and the dependent type, $Γ ⊢ φ$ for
theorems. In the latter type, $Γ$ has $\abbre{List}~\Prop$ type
and $φ$ has type $\Prop$. $\abbre{List}$ is the usual inductive type
for lists. Finally, we will denote by $φ ≡ ψ$, the syntactic
equivalence between propositions.

@@

To emulate \Metis' inference rules in type theory,
we observed that transformations performed by some rules were given
by \emph{general recursive} functions.
But algorithms based on general recursions can not be translated
directly in type theory since it is not a guarantee they terminate.
For that reason, we present in this section, a technique to avoid
such termination problems by modifying the recursive functions
to be \emph{structural recursive}.
% (see \cite{Coquand1992,Abel2002,Bove2005} for more details).

We adopt the notion of \emph{structurally smaller} from
\citeauthor{Abel2002}~\cite{Abel2002} to say, a recursive function is
structurally recursive if it calls itself with only
\emph{structurally smaller}
arguments.

@@

\begin{align}
\begin{split}
\label{eq:unshunt-definition}
&unshunt_{0}(φ) =\\
&\hspace{3mm}\begin{cases}
unshunt_{0}((φ₁ ∧ φ₂) ⇒ φ₃),
  & \text{if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
unshunt_{0}(φ₁ ⇒ φ₂) ∧ unshunt_{0}(φ₁ ⇒ φ₃),
  & \text{if }φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
φ, &\text{otherwise.}
\end{cases}
\end{split}
\end{align}

@@

Thus, we define the function
$unshunt_{1} : \Prop \to \Nat \to \Prop$
by structural recursion in its second argument in
Eq.~(\ref{eq:unshunt-structured}).

\begin{align}
\label{eq:unshunt-structured}
\begin{split}
&unshunt_{1}(φ, \zero)\hspace{3mm}= φ\\%\hspace{10cm}\\
&unshunt_{1}(φ, \suc(n)) =\\
&\hspace{3mm}\begin{cases}
unshunt_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n),
  &\text{if } φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\[2mm]
unshunt_{1}(φ₁ ⇒ φ₂, n)\\
∧\ unshunt_{1}(φ₁ ⇒ φ₃, n),
   &\text{if } φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).
\end{cases}
\end{split}
\end{align}

@@

\begin{equation*}
\label{eq:rank-definition}
rank(φ)=
\begin{cases}
0, &\text{for atomic }φ \\
rank(φ₁) + c_{¬},  &\text{if } φ ≡ \neg φ₁ \\
\max{(rank(φ₁),rank(φ₂))} + c_{\square},
  &\text{if } φ ≡ φ₁\,\square\, φ₂.
\end{cases}
\end{equation*}

@@

\begin{align}
  \begin{split}
  \label{eq:unshunt-rank-definition}
      &unshunt_{cm}(φ) =\\
      &\hspace{3mm}
        \begin{cases}
        unshunt_{cm}(φ₃) + 2, &\text{if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
        \max(unshunt_{cm}(φ₂),unshunt_{cm}(φ₃)) + 1,
          &\text{if }  φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
        0, &\text{otherwise.}
        \end{cases}
  \end{split}
\end{align}

@@

In this section, we emulate each \Metis inference rule by proving a
function with the same name, and a theorem that
proves the soundness of the function. Then, we will find
the following pattern to present the formalization for
the \Metis inference rules.

\begin{itemize}
\item \texttt{rule} is a \Metis inference rule
\item $rule$ is the emulated version function for
the \texttt{rule} inference.
\begin{align*}
\begin{split}
&rule : \Prop → \Prop → \Prop\\
&rule(φ, ψ) =
  \begin{cases}
  ψ, &\text{if }rule\text{ built }ψ\text{ by applying valid transformations to }φ\\
  φ, &\text{otherwise.}
  \end{cases}
\end{split}
\end{align*}

\begin{itemize}
  \item $φ$ is the \emph{source} formula.
  \item $ψ$ is the \emph{target} formula.
\end{itemize}
\item thm-rule is the theorem, $Γ ⊢ rule(φ, ψ)$, that proves
the soundness of the $rule$ function.
\end{itemize}

@@

To prove the correctness of the process mentioned above, we state
the Theorem~\ref{thm:thm-strip}. But first, let us prove some
lemmas for $unshunt$ function presented in Def.~(\ref{def:unshunt}),
and for $split$ function presented in Eq.~(\ref{eq:split-definition}).

@@

Let us now consider the $split$ function presented in
Eq.~(\ref{eq:split-definition}) that yields the conjunction of subgoals that implies the goal of the problem in the \Metis \TSTP derivations.

\begin{definition}[split]
\label{def:split}

\begin{equation}
\label{eq:split-definition}
\begin{split}
&split : \Prop → \Prop\\
&\mysplit(φ) =
\begin{cases}
unshunt(\mysplit(φ₁)) &\text{if }φ ≡ φ₁ ∧ φ₂\\
∧\ unshunt(φ₁ ⇒ \mysplit(φ₂)), &\\[2.5mm]
unshunt(¬ φ₁ ⇒ \mysplit(φ₂)),
  & \text{if }φ ≡ φ₁ ∨ φ₂\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(φ₂)),
  & \text{if }φ ≡ φ₁ ⇒ φ₂\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(φ₂))
  & \text{if }φ ≡ φ₁ ⇔ φ₂\\
∧\ unshunt(φ₂ ⇒ \mysplit(φ₁)),
  &\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
  & \text{if }φ ≡ ¬ (φ₁ ∧ φ₂)\\[2.5mm]
unshunt(\mysplit(¬ φ₁))
  & \text{if }φ ≡ ¬ (φ₁ ∨ φ₂)\\
∧\ unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂)),
  &\\[2.5mm]
unshunt(\mysplit(φ₁))
  & \text{if }φ ≡ ¬ (φ₁ ⇒ φ₂)\\
∧\ unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
  &\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(¬ φ₂))
  & \text{if }φ ≡ ¬ (φ₁ ⇔ φ₂)\\
∧\ unshunt(¬ φ₂ ⇒ \mysplit(φ₁)),
  &\\[2.5mm]
unshunt(\mysplit(φ₁)),
  & \text{if }φ ≡ ¬ ¬ φ₁\\
⊤,
  & \text{if }φ ≡ ¬ ⊥\\
⊥,
  & \text{if }φ ≡ ¬ ⊤.
\end{cases}
\end{split}
\end{equation}
\end{definition}

The definition above suffers termination problems since it is not
a structural recursion. Therefore, we have reformulated the equation
above to solve such a issue following the technique
described in Section~\ref{ssec:structural-recursion}. We refer to $split$
as the function after applying the process mentioned above here and below
unless otherwise stated.

@@

\begin{sketchproof}
We now proceed by induction on the structure of the
formula $φ$ by following the cases defined in
Eq.~(\ref{eq:split-definition}).
We present some cases but for a complete proof, we refer the reader to~\cite{AgdaMetis}.
\begin{itemize}
\item $(φ ≡ ¬ ¬ φ₁)$
%Using the double negation equivalence, we conclude.
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(split(¬¬φ₁,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢split(¬¬φ₁,n-1)$}
  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ ⊢ ¬¬φ₁$}
  \RightLabel{axiom.}
  \UnaryInfC{$Γ ⊢ φ$}
  \end{bprooftree}
\end{equation*}

@@

\begin{theorem}[thm-strip]
\label{thm:thm-strip}
 If $s_i$ are the subgoals of the goal $φ$, and
 $\{ Γ ⊢ s_i \}_{i}$, then $Γ ⊢ φ$.
\end{theorem}

\begin{proof}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\left\{ Γ ⊢ s_i \right\}_{i}$}
  \RightLabel{∧-intro}
  \UnaryInfC{$Γ ⊢ \bigwedge_i s_i$}

  \AxiomC{$\mysplit(φ) ≡ \bigwedge_i s_i $}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ \mysplit(φ)$}
  \RightLabel{\thmsplit.}
  \UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
\end{equation*}
\end{proof}

@@

Then, to emulate this rule, we just introduce the negation
of the subgoal by using the assume rule from the inferences in
Fig.~\ref{fig:CPL-inference-rules}.

% -------------------------------------------------------------------
\subsubsection{Splitting a Conjunct.}
\label{sssec:splitting-a-conjunct}

@@

\begin{definition}[conjunct]
\begin{align}
  \begin{split}
  &conjunct : \Prop \to \Prop \to \Prop\\
  &conjunct(φ, ψ) =
    \begin{cases}
        ψ, &\text{if }φ ≡ ψ\\
        ψ, &\text{if }φ ≡ φ₁ ∧ φ₂\text{ , }ψ ≡ conjunct(φ₁, ψ)\\
        ψ, &\text{if }φ ≡ φ₁ ∧ φ₂\text{ , }ψ ≡ conjunct(φ₂, ψ)\\
        φ, &\text{otherwise.}
      \end{cases}
  \end{split}
\end{align}
\end{definition}

@@

\subsubsection{Reordering.}
Logic equivalence between propositions is a major issue to justify
prover reasoning steps. Since we left out semantics to treat only the
syntax aspects of the propositions, our approach shows that equivalence
by converting propositions to their conjunctive normal forms and showing
equivalence by reordering the conjunctions and the inner disjunctions.
In this section, we provide some theorems to perform such reordering tasks,
the goals are the following.

\begin{itemize}
  \item Reorder the disjunction $φ = φ₁ ∨ φ₂ ∨ \cdots φₙ$ to match with
$ψ = ψ₁ ∨ ψ₂ ∨ \cdots ∨ ψₙ$, when for all $i = 1, \cdots, n$, exists at
least one $ψⱼ$ such that $φᵢ ≡ ψⱼ$ for $j = 1, \cdots, n$.
  \item Reorder a CNF formula, $φ = φ₁ ∧ φ₂ ∧ \cdots$, to match with another
CNF formula $ψ = ψ₁ ∧ ψ₂ ∧ \cdots$ when for all $i = 1, \cdots, n$, exists at
least one $ψⱼ$ such that $φᵢ ≡ ψⱼ$ for $j = 1, \cdots, n$.
\end{itemize}

First, we define the $rdisj$ function in Eq.~(\ref{eq:rdisj-def}) to convert
a disjunction in its right-associative form.

\begin{definition}[rdisj]
\label{def:rdisj}
\begin{align}
  \begin{split}
  \label{eq:rdisj-def}
    &rdisj(φ) : \Prop \to \Prop\\
    &rdisj(φ) =
    \begin{cases}
    rdisj(φ₁ ∨ (φ₂ ∨ φ₃)),  & \text{if } φ ≡ (φ₁ ∨ φ₂) ∨ φ₃\\
    φ₁ ∨ rdisj(φ₂),         & \text{if } φ ≡ φ₁ ∨ φ₂\\
    φ,                      & \text{otherwise.}
    \end{cases}
  \end{split}
\end{align}
\end{definition}

\begin{lemma}[lem-rdisj]
\label{lem:lem-rdisj}
  $Γ ⊢ φ$ then $Γ ⊢ rdisj(φ)$.
\end{lemma}
% Following a similar way to define $rdisj$, we define the $rconj$ function,
% to convert a conjunction into its right-associative form.

@@

\begin{definition}[build$_{∨}$]
\begin{equation}
  \begin{split}
  \label{eq:build-or}
  &build_{∨} : \Prop → \Prop → \Prop\\
  &build_{∨}(φ, ψ) =\\
  &\hspace{3mm}
  \begin{cases}
  ψ, &\text{if } φ ≡ ψ\\
  ψ, &\text{if } ψ ≡ ψ₁ ∨ ψ₂\text{ and } build_{∨}(φ, ψ_{i}) ≡ ψ_{i}\text{ for some }i=1, 2\\
  φ, &\text{otherwise.}
  \end{cases}
  \end{split}
\end{equation}
\end{definition}

@@

\begin{definition}[factor]
\begin{align}
\begin{split}
  \label{eq:factor-definition}
  &factor : \Prop → \Prop\\
  &factor(φ) =
  \begin{cases}
    φ₁,  &\text{if }φ ≡ φ₁ ∨ φ₂ \text{ and } φ₁ ≡ factor(φ₂)\\
    φ,   &\text{otherwise.}
  \end{cases}
\end{split}
\end{align}
\end{definition}

@@

\begin{definition}[sbuild$_{∨}$]
\begin{align}
  \begin{split}
  \label{eq:strong-build-or-definition}
    &sbuild_{∨} : \Prop → \Prop → \Prop\\
    &sbuild_{∨}(φ, ψ) =\\
    &\hspace{3mm}
    \begin{cases}
      factor(build_{∨}(φ₁, ψ) ∨ build_{∨}(φ₂,ψ)),  &\text{if }φ≡ φ₁ ∨ φ₂\\
      build_{∨}(φ, ψ),  & \text{otherwise.}
    \end{cases}
  \end{split}
\end{align}
\end{definition}

@@

\begin{definition}[reorder$_{∧}$]
  \begin{align*}
      \begin{split}
        &reorder : \Prop → \Prop \to \Prop\\
        &reorder_{∧}(φ, ψ) =
        \begin{cases}
          φ, &\text{if }φ ≡ ψ\\
          ψ₁ ∧ ψ₂, &\text{if } ψ ≡ ψ₁ ∧ ψ₂\text{, }reorder_{∧}(φ, ψ₁) ≡ ψ₁\\
                  &\text{and }reorder_{∧}(φ, ψ₂)≡ ψ₂\\
          φ,       &\text{if } ψ ≡ ψ₁ ∧ ψ₂\\
          conjunct(φ, ψ), &\text{otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

@@

\begin{definition}[conjunct$_{∨}$]
  \label{eq:conjunct-or-definition}
\begin{align*}
  \begin{split}
    &conjunct_{∨} : \Prop → \Prop \to \Prop\\
    &conjunct_{∨}(φ, ψ) =
    \begin{cases}
      ψ, &\text{if }φ ≡ ψ\\
      ψ, &\text{if }reorder_{∨}(φ, ψ) ≡ ψ\\
      ψ, &\text{if }ψ ≡ ψ₁ ∧ ψ₂,\ conjunct_{∨}(φ, ψ₁) ≡ ψ₁,\\
        &\text{and } reorder_{∨}(φ, ψ₂) ≡ ψ₂\\
      ψ, &\text{if }φ ≡ φ₁ ∧ φ₂,\ conjunct_{∨}(φ₁, ψ) ≡ ψ\\
      ψ, &\text{if }φ ≡ φ₁ ∧ φ₂,\ conjunct_{∨}(φ₂, ψ) ≡ ψ\\
      φ, &\text{otherwise.}
    \end{cases}
  \end{split}
  \end{align*}
\end{definition}

@@

\begin{definition}[reorder$_{∧∨}$]
 \begin{align*}
      \begin{split}
      &reorder : \Prop \to \Prop \to \Prop\\
      &reorder_{∧∨}(φ, ψ)=
        \begin{cases}
          ψ, &\text{if } φ≡ψ\\
          ψ, &\text{if } ψ ≡ ψ₁ ∧ ψ₂,\ reorder_{∧∨}(φ,ψ₁) ≡ ψ₁\\
             &\text{and }reorder_{∧∨}(φ, ψ₂) ≡ ψ₂\\
          φ, &\text{if }ψ ≡ ψ₁ ∧ ψ₂\\
          conjunct_{∨}(φ, ψ) &\text{otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
  \end{definition}

@@

\begin{definition}[rsol]
\label{def:rsol}
\begin{align*}
      \begin{split}
        &rsol : \Prop \to \Prop\\
        &rsol(φ) =
        \begin{cases}
          φ₂, &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∨ (φ₃ ∨ φ₄), φ₃ ≡ ¬ φ₁, φ₂ ≡ φ₄\\
          φ₂ ∨ φ₄, &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∨ (φ₃ ∨ φ₄), φ₃ ≡ ¬ φ₁\\
          φ, &\text{otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

>>>>

In the remainder part of this section, a formal description using
type theory is presented to build definitions and theorems of functions
necessary to emulate \Metis' inference rules. As notation convention,
We write the type judgments with ``$a : A$'' to denote that $a$
has $\abbre{A}$ type. Types are written using small caps.
We use \Prop type for propositions and the dependent type, $Γ ⊢ φ$ for
theorems where $Γ$ has $\abbre{List}~\Prop$ type, and $φ$ has type $\Prop$.
$\abbre{List}$ type is the usual inductive type for lists. For syntactical
equivalence, we write $φ ≡ ψ$ for $φ : \Prop$ and $ψ : \Prop$.

@@

To emulate \Metis' inference rules in type theory,
we observed that transformations performed by some rules were given
by \emph{general recursive} functions.
But algorithms based on general recursions can not be translated
directly in type theory since it is not a guarantee they terminate.
For that reason, we follow the technique described
in~\cite{Bertot2004} to avoid termination problems by modifying the
recursive
functions to be \emph{structural recursive}.
% (see \cite{Coquand1992,Abel2002,Bove2005} for more details).

This paper adopts the notion of \emph{structurally smaller} from
\citeauthor{Abel2002}~\cite{Abel2002} to say, a recursive function is
structurally recursive if it calls itself with only
\emph{structurally smaller}
arguments.

@@

\begin{align}
  \label{eq:unshunt-definition}
  \begin{split}
    &unshunt_{0} : \Prop → \Prop\\
    &unshunt_{0}(φ) =\\
    &\hspace{3mm}\begin{cases}
    unshunt_{0}((φ₁ ∧ φ₂) ⇒ φ₃),
      & \text{ if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
    unshunt_{0}(φ₁ ⇒ φ₂) ∧ unshunt_{0}(φ₁ ⇒ φ₃),
      & \text{ if }φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
    φ, &\text{ otherwise.}
    \end{cases}
  \end{split}
\end{align}

@@

Thus, we define the function
$unshunt_{1}$ by structural recursion in its second argument in
Eq.~(\ref{eq:unshunt-structured}).

\begin{align}
  \label{eq:unshunt-structured}
  \begin{split}
    &unshunt_{1} : \Prop → \Nat \to \Prop\\
    &unshunt_{1}(φ, \zero)\hspace{3mm}= φ\\%\hspace{10cm}\\
    &unshunt_{1}(φ, \suc(n)) =\\
    &\hspace{3mm}\begin{cases}
    unshunt_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n),
      &\text{ if } φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\[2mm]
    unshunt_{1}(φ₁ ⇒ φ₂, n)\\
    ∧\ unshunt_{1}(φ₁ ⇒ φ₃, n),
       &\text{ if } φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).
  \end{cases}
  \end{split}
\end{align}

@@

\begin{align*}
  \begin{split}
  % \label{eq:rank-definition}
    &rank : \Prop → \Nat\\
    &rank(φ) =
    \begin{cases}
    0, &\text{for atomic }φ \\
    rank(φ₁) + c_{¬},  &\text{ if } φ ≡ \neg φ₁ \\
    \max{(rank(φ₁),rank(φ₂))} + c_{\square},
      &\text{ if } φ ≡ φ₁\,\square\, φ₂.
    \end{cases}
  \end{split}
\end{align*}

@@

\begin{align}
  \label{eq:unshunt-rank-definition}
  \begin{split}
      &unshunt_{cm} : \Prop → \Nat\\
      &unshunt_{cm}(φ) =\\
      &\hspace{3mm}
        \begin{cases}
        unshunt_{cm}(φ₃) + 2, &\text{ if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
        \max(unshunt_{cm}(φ₂),unshunt_{cm}(φ₃)) + 1,
          &\text{ if }  φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
        0, &\text{ otherwise.}
        \end{cases}
  \end{split}
\end{align}

@@

In this section, we emulate each \Metis inference rule by proving a
function with the same name, and a theorem that
proves the soundness of the function. Then, we will find
the following pattern to present the formalization for
the \Metis inference rules. The function name written in bold print is a \Metis
inference rule in \TSTP derivation. The same function name written
in italics refers to the emulated version presents in this section.
Each emulated version states a theorem to prove the soundness of the rule.
This theorem is named by using the prefix \emph{thm}
and the name of the rule.

\begin{itemize}
\item \texttt{rule} is a $n$-ary \Metis inference rule
\item $rule$ is the emulated version function  for
the \texttt{rule} inference.
\begin{align*}
  \begin{split}
    &rule : \Prop → \Prop → \cdots → \Prop\\
    &rule(φ₁,\cdots, φₙ, ψ) =
      \begin{cases}
      ψ, &\text{ if }rule\text{ built }ψ\text{ by applying valid}\\
         &\text{ transformations or other rules to }φ\\
      φ, &\text{ otherwise.}
      \end{cases}
  \end{split}
\end{align*}

\begin{itemize}
  \item $φᵢ$ are \emph{source} formulas from previous deducted derivations.
  \item $ψ$ is the \emph{target} formula. The excepted result of \texttt{rule}.
\end{itemize}
\end{itemize}

@@

To prove the correctness of the process mentioned above, first,
we need to establish some lemmas before introduce our main goal of this part,
Theorem~\ref{thm:thm-strip}.

@@

Let us now define the $split$ function that yields the conjunction of subgoals that implies the goal of the problem in the \Metis \TSTP derivations.

\begin{definition}[split]
\label{def:split}
  \begin{equation*}
    \begin{split}
    &split : \Prop → \Prop\\
    &\mysplit(φ) =
      \begin{cases}
      unshunt(\mysplit(φ₁)) &\text{ if }φ ≡ φ₁ ∧ φ₂\\
      ∧\ unshunt(φ₁ ⇒ \mysplit(φ₂)), &\\[2.5mm]
      unshunt(¬ φ₁ ⇒ \mysplit(φ₂)),
        & \text{ if }φ ≡ φ₁ ∨ φ₂\\[2.5mm]
      unshunt(φ₁ ⇒ \mysplit(φ₂)),
        & \text{ if }φ ≡ φ₁ ⇒ φ₂\\[2.5mm]
      unshunt(φ₁ ⇒ \mysplit(φ₂))
        & \text{ if }φ ≡ φ₁ ⇔ φ₂\\
      ∧\ unshunt(φ₂ ⇒ \mysplit(φ₁)),
        &\\[2.5mm]
      unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
        & \text{ if }φ ≡ ¬ (φ₁ ∧ φ₂)\\[2.5mm]
      unshunt(\mysplit(¬ φ₁))
        & \text{ if }φ ≡ ¬ (φ₁ ∨ φ₂)\\
      ∧\ unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂)),
        &\\[2.5mm]
      unshunt(\mysplit(φ₁))
        & \text{ if }φ ≡ ¬ (φ₁ ⇒ φ₂)\\
      ∧\ unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
        &\\[2.5mm]
      unshunt(φ₁ ⇒ \mysplit(¬ φ₂))
        & \text{ if }φ ≡ ¬ (φ₁ ⇔ φ₂)\\
      ∧\ unshunt(¬ φ₂ ⇒ \mysplit(φ₁)),
        &\\[2.5mm]
      unshunt(\mysplit(φ₁)),
        & \text{ if }φ ≡ ¬ ¬ φ₁\\
      ⊤,
        & \text{ if }φ ≡ ¬ ⊥\\
      ⊥,
        & \text{ if }φ ≡ ¬ ⊤.
      \end{cases}
    \end{split}
  \end{equation*}
\end{definition}

The definition above suffers termination problems since it is not
a structural recursion. Therefore, we have reformulated the equation
above to solve such an issue following the technique
described in Section~\ref{ssec:structural-recursion}. We refer to $split$
as the function after applying the process mentioned above here and below
unless otherwise stated.

@@

\begin{sketchproof}
The proof is by induction on the structure of the
formula $φ$ by following the cases in Def.~\ref{def:split}.
We present some cases but for a complete proof, we refer the reader to~\cite{AgdaMetis}.
\begin{itemize}
\item $(φ ≡ ¬ ¬ φ₁)$
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(split(¬¬φ₁,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢split(¬¬φ₁,n-1)$}
  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ ⊢ ¬¬φ₁$}
  \RightLabel{axiom.}
  \UnaryInfC{$Γ ⊢ φ$}
  \end{bprooftree}
\end{equation*}

@@

We can now formulate our main result of this part that
justifies the stripping strategy of \Metis to prove goals.

\begin{theorem}[thm-strip]
\label{thm:thm-strip}
 If $s_i$ are the subgoals of the goal $φ$ for $i=1,\cdots,n$, and
 $\{ Γ ⊢ s_i \}_{i =1}^n$, then $Γ ⊢ φ$.
\end{theorem}

\begin{proof}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\left\{ Γ ⊢ s_i \right\}_{i=1}^n$}
  \RightLabel{∧-intro}
  \UnaryInfC{$Γ ⊢ \bigwedge_{i=1}^n s_i$}

  \AxiomC{$\mysplit(φ) ≡ \bigwedge_{i=1}^n s_i $}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ \mysplit(φ)$}
  \RightLabel{\thmsplit.}
  \UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
\end{equation*}
\end{proof}

@@

Then, to emulate this rule, we just introduce the negation
of the subgoal by using $assume$ rule from the inferences in
Fig.~\ref{fig:CPL-inference-rules}.

\subsubsection{Splitting a Conjunct.}
\label{sssec:splitting-a-conjunct}

@@

\begin{definition}[conjunct]
\begin{align}
  \begin{split}
  &conjunct : \Prop \to \Prop \to \Prop\\
  &conjunct(φ, ψ) =
    \begin{cases}
        ψ, &\text{ if }φ ≡ ψ\\
        ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂\text{ , }ψ ≡ conjunct(φ₁, ψ)\\
        ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂\text{ , }ψ ≡ conjunct(φ₂, ψ)\\
        φ, &\text{ otherwise.}
      \end{cases}
  \end{split}
\end{align}
\end{definition}

@@

\subsubsection{Reordering.}
Logic equivalence between propositions is a major issue to justify
prover reasoning steps. Since we left out semantics to treat only the
syntax aspects of the propositions, our approach shows equivalence
by converting propositions to their conjunctive normal forms.
But also, reordering the conjunctions and the inner disjunctions into the
formulas to match them.
In this section, we provide some theorems to perform such reordering tasks,
the goals are the following.

\begin{itemize}
  \item Reorder a disjunction $φ = φ₁ ∨ φ₂ ∨ \cdots φₙ$ to match with
$ψ = ψ₁ ∨ ψ₂ ∨ \cdots ∨ ψₙ$, when for all $i = 1, \cdots, n$, exists at
least one $ψⱼ$ such that $φᵢ ≡ ψⱼ$ for $j = 1, \cdots, n$.
  \item Reorder a CNF formula, $φ = φ₁ ∧ φ₂ ∧ \cdots$, to match with another
CNF formula $ψ = ψ₁ ∧ ψ₂ ∧ \cdots$ when for all $i = 1, \cdots, n$, exists at
least one $ψⱼ$ such that $φᵢ ≡ ψⱼ$ for $j = 1, \cdots, n$.
\end{itemize}

First, we define the $rdisj$ function to convert
a disjunction in its right-associative form.
The $rconj$ function converts conjunction to its right-associative form is not
presented here, but is defined in a similar way (see its implementation
in~\cite{AgdaMetis}).

\begin{definition}[rdisj]
\label{def:rdisj}
  \begin{align*}
    \begin{split}
    \label{eq:rdisj-def}
      &rdisj(φ) : \Prop \to \Prop\\
      &rdisj(φ) =
      \begin{cases}
      rdisj(φ₁ ∨ (φ₂ ∨ φ₃)),  & \text{ if } φ ≡ (φ₁ ∨ φ₂) ∨ φ₃\\
      φ₁ ∨ rdisj(φ₂),         & \text{ if } φ ≡ φ₁ ∨ φ₂\\
      φ,                      & \text{ otherwise.}
      \end{cases}
    \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-rdisj]
\label{lem:lem-rdisj}
  $Γ ⊢ φ$ then $Γ ⊢ rdisj(φ)$.
\end{lemma}

@@

\begin{definition}[build$_{∨}$]
\begin{equation}
  \begin{split}
  \label{eq:build-or}
  &build_{∨} : \Prop → \Prop → \Prop\\
  &build_{∨}(φ, ψ) =\\
  &\hspace{3mm}
  \begin{cases}
  ψ, &\text{ if } φ ≡ ψ\\
  ψ, &\text{ if } ψ ≡ ψ₁ ∨ ψ₂\text{, } build_{∨}(φ, ψ_{i}) ≡ ψ_{i}\text{ for some }i=1, 2\\
  φ, &\text{ otherwise.}
  \end{cases}
  \end{split}
\end{equation}
\end{definition}

@@

\begin{definition}[factor]
\begin{align}
\begin{split}
  \label{eq:factor-definition}
  &factor : \Prop → \Prop\\
  &factor(φ) =
  \begin{cases}
    φ₁,  &\text{ if }φ ≡ φ₁ ∨ φ₂ \text{, } φ₁ ≡ factor(φ₂)\\
    φ,   &\text{ otherwise.}
  \end{cases}
\end{split}
\end{align}
\end{definition}

@@

\begin{definition}[sbuild$_{∨}$]
\begin{align}
  \begin{split}
  \label{eq:strong-build-or-definition}
    &sbuild_{∨} : \Prop → \Prop → \Prop\\
    &sbuild_{∨}(φ, ψ) =\\
    &\hspace{3mm}
    \begin{cases}
      factor(build_{∨}(φ₁, ψ) ∨ build_{∨}(φ₂,ψ)),  &\text{ if }φ≡ φ₁ ∨ φ₂\\
      build_{∨}(φ, ψ),  & \text{ otherwise.}
    \end{cases}
  \end{split}
\end{align}
\end{definition}

@@

\begin{definition}[reorder$_{∧}$]
  \begin{align*}
      \begin{split}
        &reorder : \Prop → \Prop \to \Prop\\
        &reorder_{∧}(φ, ψ) =
        \begin{cases}
          φ, &\text{ if }φ ≡ ψ\\
          ψ₁ ∧ ψ₂, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\text{, }reorder_{∧}(φ, ψ₁) ≡ ψ₁\\
                  &\text{and }reorder_{∧}(φ, ψ₂)≡ ψ₂\\
          φ,       &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\\
          conjunct(φ, ψ), &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

@@

\begin{definition}[conjunct$_{∨}$]
  \label{eq:conjunct-or-definition}
\begin{align*}
  \begin{split}
    &conjunct_{∨} : \Prop → \Prop \to \Prop\\
    &conjunct_{∨}(φ, ψ) =
    \begin{cases}
      ψ, &\text{ if }φ ≡ ψ\\
      ψ, &\text{ if }reorder_{∨}(φ, ψ) ≡ ψ\\
      ψ, &\text{ if }ψ ≡ ψ₁ ∧ ψ₂,\ conjunct_{∨}(φ, ψ₁) ≡ ψ₁,\\
        &\text{and } reorder_{∨}(φ, ψ₂) ≡ ψ₂\\
      ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂,\ conjunct_{∨}(φ₁, ψ) ≡ ψ\\
      ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂,\ conjunct_{∨}(φ₂, ψ) ≡ ψ\\
      φ, &\text{ otherwise.}
    \end{cases}
  \end{split}
  \end{align*}
\end{definition}

@@

\begin{definition}[reorder$_{∧∨}$]
 \begin{align*}
      \begin{split}
      &reorder : \Prop \to \Prop \to \Prop\\
      &reorder_{∧∨}(φ, ψ)=
        \begin{cases}
          ψ, &\text{ if } φ≡ψ\\
          ψ, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂,\ reorder_{∧∨}(φ,ψ₁) ≡ ψ₁\\
             &\text{and }reorder_{∧∨}(φ, ψ₂) ≡ ψ₂\\
          φ, &\text{ if }ψ ≡ ψ₁ ∧ ψ₂\\
          conjunct_{∨}(φ, ψ) &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
  \end{definition}

@@

\begin{definition}[rsol]
\label{def:rsol}
\begin{align*}
      \begin{split}
        &rsol : \Prop \to \Prop\\
        &rsol(φ) =
        \begin{cases}
          φ₂, &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∨ (φ₃ ∨ φ₄), φ₃ ≡ ¬ φ₁, φ₂ ≡ φ₄\\
          φ₂ ∨ φ₄, &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∨ (φ₃ ∨ φ₄), φ₃ ≡ ¬ φ₁\\
          φ, &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}
